---
title: "Untitled"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(stringr)
library(vars)
```


```{r}
data1 = read.csv('/Users/nimesh/Documents/TimeSeries/train_project.csv')
test = read.csv('/Users/nimesh/Documents/TimeSeries/test_project.csv')

data =  data1[1:264,]
val  = data1[265:288,]
data = ts(data, start = c(1987,1), frequency = 12)
```

```{r}
#for (i in seq(1,nrow(data),1))
#{
#data[i,'Month'] = paste( str_sub(data[i,'Month'],-6,-5),str_sub(data[i,'Month'],-4,-1), sep='/')
#}
#data['Month']
```
```{r}
plot(ts(data[,2]), ylab = 'Unemployment_rate')
```
```{r}
plot(ts(data[,3]), ylab = 'Population')
```

```{r}
plot(log(data[,4]), ylab = 'Bankruptcy_rate')
```

The plot shows that the variance of the data changes with time. So we wapply log transformation on the data.
```{r}
plot(ts(data[,5]), ylab = 'House_Price_Index')
```

```{r}
br = log(data[,4])
plot(br)
```

After applying the log transformation, the data becomes homoscedastic.

```{r}
adf.test(br)
br_1 <- diff(br)
```

```{r}
adf.test(br_1)
```
The p-value is less than 0.05. The data is stationary.

```{r}
br_2 <- diff(br_1, lag = 12)
```

```{r}

acf(br_2, lag.max = 72)
```
q=1,3,4
Q=2,3,4
```{r}
pacf(br_2, lag.max = 48)
```

p =1/2 ,P = 0
```{r}
valid_ts <- ts(data = val$Bankruptcy_Rate, start = c(2009,1), frequency = 12)

print(paste("Model p,q,P,Q,d;    RMSE;          Variance;            AIC; LogLik"))
#permutations(n=3,r=2,v=x,repeats.allowed=T)
grid <-expand.grid(0:5, 0:5,0:5, 0:5, :5)
grid <- as.data.frame(grid)
names(grid) <- c("Var1", "Var2", "Var3", "Var4", "Var5")

for (i in 1:nrow(grid)){
p <- grid$Var1[i] 
q <- grid$Var2[i]
P <- grid$Var3[i]
Q <- grid$Var4[i]
d <- grid$Var5[i]
  
m <- tryCatch(arima(x = log_br_ts, order = c(p,d,q), seasonal = list(order = c(P,1,Q), period = 12), method = "CSS", optim.control = list(maxit = 1000)),
              error = function(e) {print(paste("gave error"))})

m.rmse <- tryCatch(rmse(exp(forecast(m, h = 24)$mean), valid_ts),
                   error = function(e) {print(paste("gave error"))})

print(paste("p" = p, "q" = q, "P" = P, "Q" = Q,"d" = d, " = ",  m.rmse, m$sigma2, 
            m$aic, m$loglik))

}
```
```{r}
for  (p  in  seq(0,  5)){
  for  (q  in  seq(0,  5)){
    for  (P  in  seq(0,  5)){
      for  (Q  in  seq(0,  5)){
        model  <-  arima(data[,4], 
                       order  =  c(p,  1,  q), 
                       seasonal  =  list(order  =  c(P,  1,  Q),  period  =  12), 
                       method  =  "CSS", 
                       xreg  =  data.frame(data[,4], data['Unemployment_rate']))
        
      
          
        f  <-  forecast(model,  h  =  24,  level  =  0.95,xreg = data.frame(val['Population'],val['Unemployment_Rate']))
        rmse_c  <-  sqrt(mean((val['Bankruptcy_Rate']  -  exp(f$fitted))  ^  2))
        

       print(rmse_c)
      }
    } 
  }
}

```

```{r}
m1 <- arima(br_2, order = c(2, 0, 1), seasonal = list(order = c(0, 0, 2), period = 12), method = "ML")
m2 <- arima(br_2, order = c(2,0,1), seasonal = list(order = c(0, 0, 3), period = 12), method = "ML")
m3 <- arima(br, order = c(2,1,3), seasonal = list(order = c(0, 1, 3), period = 12), method = "ML")
#m4 <- arima(data, order = c(4,1,3), seasonal = list(order = c(1,1,2), period = 12), method = "ML")
```
```{r}
VARselect(y = data.frame(Bev, Food, Indust))
```

```{r}
D <- -2*(m1$loglik - m2$loglik)
pval <- 1-pchisq(D,1)
print(c("Test Statistic:",round(D, 4),"P-value:", round(pval, 4)))
```
```{r}
D <- -2*(m1$loglik - m3$loglik)
pval <- 1-pchisq(D,2)
print(c("Test Statistic:",round(D, 4),"P-value:", round(pval, 4)))
```

```{r}
pred <- forecast(object = m3, h = 24, level = 0.95)
par(mfrow=c(1,1))
plot(exp(br), xlim=c(1987,2014), ylim=c(0,0.09)) #plotting the data
points(ts(val[,4], start = c(2009,1), frequency = 12), type='l')
abline(v=2003,col='blue',lty=2) # adding a vertical line at the point where prediction starts
points(exp(pred$mean),type='l',col='red')# plotting the predict
points(exp(pred$lower),type='l', col='green') # plotting lower limit of the prediction interval
points(exp(pred$upper),type='l', col='green') # plotting upper limit of the  prediction interval
legend(1987, 23, legend = c("observed values", "prediction interval", "prediction"),
       col = c("black", "green", "red"), lty = 1, cex = 0.8)
```

```{r}
m4 <- arima(br,order  =  c(2,  1,  0), 
             seasonal  =  list(order  =  c(0,  0,  2),  period  =  12),
             method  =  "ML",  xreg  =  data)
m5 <- arima(br,order  =  c(4,  1,  5), 
             seasonal  =  list(order  =  c(0,  0,  1),  period  =  12),
             method  =  "ML",  xreg  =  data)
pred <- forecast(object = m4, h = 24, level = 0.95,  xreg  =  val)
par(mfrow=c(1,1))
plot(exp(br), xlim=c(1987,2014), ylim=c(0,0.06), ylab="Bankruptcy rate") #plotting the data
points(ts(val[,4], start = c(2009,1), frequency = 12), type='l')
abline(v=2003,col='blue',lty=2) # adding a vertical line at the point where prediction starts
points(exp(pred$mean),type='l',col='red')# plotting the predict
points(exp(pred$lower),type='l', col='blue') # plotting lower limit of the prediction interval
points(exp(pred$upper),type='l', col='green') # plotting upper limit of the  prediction interval
legend(1987, 23, legend = c("observed values", "prediction interval", "prediction"),
       col = c("black", "green", "red"), lty = 1, cex = 0.8)
sqrt(sum((exp(pred$mean)-val$Bankruptcy_Rate)**2)/length(test))
```








```{r}
VARselect(y = data.frame(data[,1],data[,2],data[,3], data[,4]))
```
```{r}
par(mar=c(1,1,1,1))
m.var <- VAR(y = data.frame( data[,3], data[,4]), p = 10,season=12)
#plot(m.var)

# Let's now do some forecasting with this model
pred <- predict(m.var, n.ahead = 24, ci = 0.95)
plot(pred)
pred$fcst$data...4. <- as.data.frame(pred$fcst$data...4.)
rmse.var <- sqrt(mean((pred$fcst$data...4.$fcst - val$Bankruptcy_Rate)^2))
rmse.var
  
```


##Holt-Winter approach:

Since time series of response variable(Bankruptcy_rate) includes both trend and seasonality we will use triple exponential smoothing for modeling using Holt-Winters approach. 
We have used Jan,1987 to Dec, 2008 data as trainign set and Jan,2009 to Dec, 2010 as test set.
```{r}
data1 = read.csv('/Users/nimesh/Documents/TimeSeries/train_project.csv')
test = read.csv('/Users/nimesh/Documents/TimeSeries/test_project.csv')

data =  data1[1:264,]
val  = data1[265:288,]
data = ts(data, start = c(1987,1), frequency = 12)
n = list()
p = list()
q = list()
r = list()
cnt = 1
for  (a  in  seq(0.1, 0.8 ,0.1)){
  for  (b  in  seq(0.1, 0.8,0.1)){
    for  (c  in  seq(0.1, 0.8 , 0.1)){
      
      hwfit <- HoltWinters(log(data[,4]),alpha=a,beta=b,gamma=c,seasonal = "multiplicative" )
      hwpred <- forecast(hwfit, h = 24, level = 0.95)
      rmse_hw <- sqrt(mean((exp(hwpred$mean)-val$Bankruptcy_Rate)**2))
      n[cnt] <- rmse_hw
      cnt <- cnt+1
      p[cnt] <- a
      q[cnt] <-b
      r[cnt] <- c
    }
  }
}
print(paste('The lowest value of rmse is',n[which.min(n)],':')  )   
```


The lowest value of rmse we achieved using Holt Winters approach of Triple exponential smoothing is `n[which.min(n)]` for alpha = `p[which.min(n)]`  , beta = `q[which.min(n)]` and gamma = `r[which.min(n)]`

##Elastic Net approach

Lets now use Elastic net for predciting the Bankruptcy rate for our validation set and check the rmse value.


```{r}
data <- read.csv('/Users/nimesh/Documents/TimeSeries/train_project.csv', sep = ',')
set.seed(1) 
train <- data[1:264, ]
val <- data[265:288, ]
y.train <- train$Bankruptcy_Rate
y.val <- val$Bankruptcy_Rate

grid.lambda <- 10^seq(10, -2, length = 100)
traintest=rbind(train,val)
traintest=traintest[,-1]
X = model.matrix(Bankruptcy_Rate ~., data = traintest)
EN.model.train <- glmnet(X[1:nrow(train),], exp(y.train), alpha = 0.5, lambda = grid.lambda)

set.seed(1) #for reproducability
cv.out <- cv.glmnet(X[1:nrow(train),], exp(y.train), alpha = 0.5)
#Find the best lambda value
best.lambda <- cv.out$lambda.min

plot(cv.out)
abline(v = log(best.lambda), col = "blue", lwd = 2)
model2 <-  glmnet(X[1:nrow(train),], log(y.train), alpha = 0.5, lambda = best.lambda)

pred <- predict(model2, lambda = best.lambda, newx = X[-(1:nrow(train)),])
rmse_2 <- sqrt(mean((exp(EN.pred) - y.val)^2)) 
print(paste('The rmse for elastic net model is ', rmse_2, ':'))
```


The value of lambda which we used for fitting our model is `best.lambda` and the rmse we achieved is `rmse.EN`